scRNAseq

### 2018-02-14 datasets/scrnaseq/scRNAseq-latest.Rmd
updated scRNAseq-latest.Rmd

```{r, workspace-setup, eval=FALSE}
#installation
# source("https://bioconductor.org/biocLite.R")
# biocLite("swfdr")
# biocLite("IHW")
# biocLite("qvalue")
# install.packages("ashr")
# install.packages("fdrtool")
# biocLite("edgeR")
# source("https://bioconductor.org/biocLite.R")
# biocLite("scran")
#  Samseq installation
# source("https://bioconductor.org/biocLite.R")
# biocLite("impute") 
# install.packages("samr")
# source("https://bioconductor.org/biocLite.R")
#biocLite("MultiAssayExperiment")
#source("https://bioconductor.org/biocLite.R")
#biocLite("SummarizedExperiment")
#biocLite("BiocGenerics")
#biocLite("S4Vectors")
#biocLite("BiocParallel")
#install_github("areyesq89/SummarizedBenchmark")
#biocLite("EBSeq")
#biocLite("scDD")
#source("https://bioconductor.org/biocLite.R")
#biocLite("SingleCellExperiment")
#source("https://bioconductor.org/biocLite.R")
#biocLite("MAST")

@@

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)

@@

```{r, MAST}
names(groups)=colnames(g_cdata_sub)
compute_MASTtpmde(g_data_sub,groups,"mouseplate_masttpm.rds")

@@

```{r,cov,eval=TRUE}
sourceDirectory("/Users/subraman/Documents/research/fdr_jc/src/benchmark-fdr/datasets/R")
source("/Users/subraman/Documents/research/fdr_jc/src/benchmark-fdr/datasets/R/strat_hist.R")
h_cov=compute_covariates_scrnaseq("h_g_nofilter_subdata.rds")
wilcox=readRDS("h_data_wilcox_p.rds")
h_wilcox_cov=data.frame(genes=names(wilcox),pvalue=wilcox)
h_wilcox_cov=h_wilcox_cov%>%left_join(h_cov,by=c("genes"))
dat=data.frame(pvalue=h_wilcox_cov$pvalue,covariate=h_wilcox_cov$median_exp)
p=strat_hist(h_wilcox_cov,pvalue="pvalue",covariate="median_exp",numQ=5)

hmast=readRDS("humanplate_masttpm.rds")
h_mast_cov=
h_scdd_cov=
p=strat_hist()
scdd_m=metadata(scDatEx)
scdd_p_df=data.frame(genes=scdd_m$Genes$gene,scdd_pval=matrix(scdd_m$Genes$nonzero.pvalue))
wilcox_p_df=data.frame(genes=names(wilcox_p),wilcox_pval=matrix(wilcox_p))

@@

## Benchmarking covariate adjusted FDR tools for application

```{r, benchmarking, eval=FALSE}
g_cdata_sub_se=apply(g_cdata_sub,1,function(x) sd(x)/sqrt(length(x))) #standard error
g_cdata_sub_es=apply(g_cdata_sub,1,function(x) mean(x[1:26])-mean(x[27:52])) #difference of means

b_dat=data.frame(pval=dat$wilcox_pval, ind_covariate=dat$mean_exp,effect_size=matrix(g_cdata_sub_es),SE=matrix(g_cdata_sub_se))
b_dat=b_dat%>%mutate(zscore = effect_size/SE)
saveRDS(b_dat,"wilcox_meanexp.RDS")
cutoffs <- c(0.01, 0.025, 0.05, 0.10, 0.20)
resfile <- paste0(getwd(), "/scrnaseqhuman_wilcox_meanexp_results_",
                  nrow(b_dat), ".RData")
library(devtools)
source("https://install-github.me/mangothecat/callr")
source("https://install-github.me/r-lib/remotes")
remotes::install_github('jgscott/FDRreg', subdir="R_pkg/")

source("benchmark-fdr/datasets/R/evaluation-helpers_mod.R")
if (!file.exists(resfile)){
  t1 <- proc.time()
  b_dat_results <- run_benchmarks_mod(dat=b_dat, alphas=cutoffs, pvals=TRUE)
  save(b_dat_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(b_dat), " genes"))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff

pdf("scrnaseq_wilcox_results.pdf",width=20,height=10)
p <- ggplot(b_dat_results$stats, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("scRNAseq with Mean gene expression Covariate, wilcox") +
  theme_classic()
print(p)

# log-scale y axis
print(p + scale_y_continuous(trans="log2") + 
        ylab("Number of rejections (log2 scale)"))
dev.off()

b_dat=data.frame(pval=dat$scdd_pval, ind_covariate=dat$mean_exp,effect_size=matrix(g_cdata_sub_es),SE=matrix(g_cdata_sub_se))
b_dat=b_dat%>%mutate(zscore = effect_size/SE)
saveRDS(b_dat,"wilcox_meanexp.RDS")

resfile <- paste0(getwd(), "/scrnaseqhuman_scdd_meanexp_results_",
                  nrow(b_dat), ".RData")
if (!file.exists(resfile)){
  t1 <- proc.time()
  b_dat_results <- run_benchmarks_mod(dat=b_dat%>%filter(!is.na(pval)), alphas=cutoffs, pvals=TRUE)
  save(b_dat_results, file=resfile)
  message(paste0("Took ", round((proc.time()-t1)[3],1),
                 " seconds for ", nrow(b_dat), " genes"))
}else{
  load(resfile)
}

# plot N rejects vs FDR cutoff

pdf("scrnaseq_scdd_results.pdf",width=20,height=10)
p <- ggplot(b_dat_results$stats, aes(alpha, n_rejects, color=method)) +
  geom_point() +
  geom_line() +
  xlab("FDR cutoff") + ylab("Number of rejections") +
  ggtitle("scRNAseq with Mean gene expression Covariate, scdd") +
  theme_classic()
print(p)

# log-scale y axis
print(p + scale_y_continuous(trans="log2") + 
        ylab("Number of rejections (log2 scale)"))
dev.off()

>>>>

```{r, workspace-setup, eval=FALSE}
#installation
# source("https://bioconductor.org/biocLite.R")
# biocLite("swfdr")
# biocLite("IHW")
# biocLite("qvalue")
# install.packages("ashr")
# install.packages("fdrtool")
# biocLite("edgeR")
# source("https://bioconductor.org/biocLite.R")
# biocLite("scran")
#  Samseq installation
# source("https://bioconductor.org/biocLite.R")
# biocLite("impute") 
# install.packages("samr")
# source("https://bioconductor.org/biocLite.R")
#biocLite("MultiAssayExperiment")
#source("https://bioconductor.org/biocLite.R")
#biocLite("SummarizedExperiment")
#biocLite("BiocGenerics")
#biocLite("S4Vectors")
#biocLite("BiocParallel")
#install_github("areyesq89/SummarizedBenchmark")
#biocLite("EBSeq")
#biocLite("scDD")
#source("https://bioconductor.org/biocLite.R")
#biocLite("SingleCellExperiment")
#source("https://bioconductor.org/biocLite.R")
#biocLite("MAST")
#install.packages("FDRreg")

@@

# benchmark methods
library(IHW)
library(ashr)
library(qvalue)
library(swfdr)
library(fdrtool)
library(FDRreg)

@@

```{r, MAST}
g_data_sub=readRDS("m_g_nofilter_subdata.rds")
groups=c(rep(1,50),rep(2,50))
names(groups)=colnames(g_cdata_sub)
compute_MASTtpmde(g_data_sub,groups,"mouseplate_masttpm.rds")

@@

```{r,cov,eval=TRUE}
sourceDirectory("/Users/subraman/Documents/research/fdr_jc/src/benchmark-fdr/datasets/R")
source("/Users/subraman/Documents/research/fdr_jc/src/benchmark-fdr/datasets/R/strat_hist.R")

h_cov=compute_covariates_scrnaseq("h_g_nofilter_subdata.rds")

wilcox=readRDS("h_data_wilcox_p.rds")
h_wilcox_cov=data.frame(genes=names(wilcox),pval=wilcox)
h_wilcox_cov=h_wilcox_cov%>%left_join(h_cov,by=c("genes"))

plot_cov_diagnosis(h_wilcox_cov,"humanplate_wilcox")
plot_cov_diagnosis(h_wilcox_cov%>%subset(detection_rate>=0.25),"humanplate_wilcox_filter")

hmast=readRDS("humanplate_masttpm.rds")
h_mast_cov=data.frame(genes=rownames(hmast),pval=hmast$pval)
h_mast_cov=h_mast_cov%>%left_join(h_cov,by=c("genes"))
plot_cov_diagnosis(h_mast_cov,"humanplate_mast")
plot_cov_diagnosis(h_mast_cov%>%subset(detection_rate>=0.25),"humanplate_mast_filter")

h_scdd=readRDS("humanplate_scDatEx_ks.rds")
h_scdd_cov=data.frame(genes=rownames(metadata(h_scdd)$Genes),pval=metadata(h_scdd)$Genes$nonzero.pvalue)
h_scdd_cov$genes=h_cov$genes #scdd changes gene names by appending numeric indices 
h_scdd_cov=h_scdd_cov%>%subset(!is.na(pval))
h_scdd_cov=h_scdd_cov%>%left_join(h_cov,by=c("genes"))
plot_cov_diagnosis(h_scdd_cov,"humanplate_scdd")
plot_cov_diagnosis(h_scdd_cov%>%subset(detection_rate>=0.25),"humanplate_scdd_filter")

m_cov=compute_covariates_scrnaseq("m_g_nofilter_subdata.rds")

wilcox=readRDS("./mouse_plate/m_data_wilcox_p.rds")
m_wilcox_cov=data.frame(genes=names(wilcox),pval=wilcox)
m_wilcox_cov=m_wilcox_cov%>%left_join(m_cov,by=c("genes"))

plot_cov_diagnosis(m_wilcox_cov,"mouseplate_wilcox")

m_wilcox_cov=m_wilcox_cov%>%subset(detection_rate>=0.25)
plot_cov_diagnosis(m_wilcox_cov,"mouseplate_wilcox_filter")

mmast=readRDS("./mouse_plate/mouseplate_masttpm.rds")
m_mast_cov=data.frame(genes=rownames(mmast),pval=mmast$pval)
m_mast_cov=m_mast_cov%>%left_join(m_cov,by=c("genes"))
plot_cov_diagnosis(m_mast_cov,"mouseplate_mast")

m_mast_cov=m_mast_cov%>%subset(detection_rate>=0.25)
plot_cov_diagnosis(m_mast_cov,"mouseplate_mast_filter")

m_scdd=readRDS("mouseplate_scDatEx_ks.rds")
m_scdd_cov=data.frame(genes=rownames(metadata(m_scdd)$Genes),pval=metadata(m_scdd)$Genes$nonzero.pvalue)
m_scdd_cov$genes=m_cov$genes
m_scdd_cov=m_scdd_cov%>%subset(!is.na(pval))
m_scdd_cov=m_scdd_cov%>%left_join(m_cov,by=c("genes"))
plot_cov_diagnosis(m_scdd_cov,"mouseplate_scdd")
plot_cov_diagnosis(m_scdd_cov%>%subset(detection_rate>=0.25),"mouseplate_scdd_filter")

@@
